---
title: "Assignment1"
author: "PingYi"
date: "21 March 2019"
output:
  word_document: default
  html_document:
  df_print: paged
  pdf_document: default
---

Q1: Use seq(), rep() and/or other commonly-used operators/functions and/or the recycling rule, but definitely not c(), nor any explicit loop, to create the following sequences.

```{r}
# 0 15 30 45 60 ... 315 330 345
seq(0, to = 345, by = 15)

```
```{r}
# 1 2 4 8 16 32 64 128 256
2^(0:8)
```

```{r}
# 1 1 1 1 2 2 2 3 3 4 
rep(1:4,4:1)
```

```{r}
# 10 11 12 13 20 21 22 23 30 31 32 33 40 41 42 43
rep(1:4, each = 4) * 10 + 0:3 
```

```{r}
# "x^1/1 + -x^2/2 + x^3/3 + -x^4/4 + x^5/5 + -x^6/6"
paste(substring(rep("-", 6), first = 1, last = 0:1), "x^", 1:6, "/", 1:6,sep = "", collapse =" + ")
```



Q2: This question makes use of data on 10,000 electric scooter trips in Austin, Texas.1 Down-load the file "AustinDockless.csv" from Canvas and run the following code to load the data into an R data frame.
The only variables we will use in this assignment are:
  . Trip.Duration (in seconds),
  . Trip.Distance (in metres),
  . the Hour during which the trip occurred,
  . Origin.Cell.ID, and Destination.Cell.ID (the city has been divided up into hexagonal "cells" each with an edge length of 500 feet).

Write R expressions to perform each of the tasks described below.


```{r}
dockless <- read.csv("AustinDockless.csv")

# Find the length of the ten longest trips.
sort(dockless[["Trip.Distance"]], decreasing = TRUE)[1:10]


```
```{r}
# Find the start and end Cell IDs for trips longer than 19km
dockless[dockless$Trip.Distance > 19000, c("Origin.Cell.ID", "Destination.Cell.ID")]

```

```{r}
# Find the longest trip (distance) between midnight and 8am.

# extract hours and am/pm indicator from Start.Time, valid Start time: 0/12am ~ 8am
startHour = as.numeric(substring(dockless$Start.Time, 12, 13))
startInd = substring(dockless$Start.Time, 21, 22)
ifStartValid = (startHour < 8 | startHour == 12) & startInd == "AM"

# extract hours and am/pm indicator from End.Time, valid End time: 0/12am ~ 8am
endHour = as.numeric(substring(dockless$End.Time, 12, 13))
endInd =  substring(dockless$End.Time, 21, 22)
ifEndValid = (endHour < 8 | endHour == 12) & endInd == "AM"

#longest trip during this period
max(dockless[ifStartValid & ifEndValid,]$Trip.Distance)

```

```{r}
# Find the average speed and the distance covered for the top ten fastest trips (trips with highest average speed), where average speed is calculated as distance divided by duration.

# construct a new data frame containing the calculated column: Speed
Speed = dockless$Trip.Distance/dockless$Trip.Duration
tempDF = cbind(dockless, Speed)

# Odering the new data frame by column Speed
tempDF = tempDF[order(tempDF[,"Speed"], decreasing = TRUE), ]

# list the top 10 fastest trips
tempDF[1:10, c("Trip.Distance", "Speed")]

```

Q3 This question also works with the electronic scooter data. We will just focus on the distance variable Trip.Distance.

```{r}
# Calculate the mean and standard deviation of the distance data.
distMean = mean(dockless$Trip.Distance)
distSD = sd(dockless$Trip.Distance)
distMean
distSD

```

```{r}
# Calculate the median and upper and lower quartiles for the distance data (HINT: use the quantile() function).
distQuartiles = quantile(dockless$Trip.Distance)[2:4]
distQuartiles
```

```{r}
# Generate 10,000 random values from a Normal distribution (HINT: use the rnorm() function) with the same mean and standard deviation as the distance data and calculate the median and upper and lower quartiles of these random values.
set.seed(1234)
randValues = rnorm(10000, distMean, distSD)
randQuartiles = quantile(randValues)[2:4]
randQuartiles

```

```{r}
# Generate 1,000,000 random values from a Normal distribution with the same mean and standard deviation as the distance data and, dealing with each consecutive subset of 10,000 values, calculate 100 upper quartiles (one for each of the 100 subsets of 10,000 values)
set.seed(1234)
randValuesT = rnorm(1000000, distMean, distSD)

# each column contains each consecutive subset of 10,000 values, 100 columns in total
randMatrix = matrix(randValuesT, ncol = 100)

# get upper quartiles for each column
uppers = apply(randMatrix, 2, quantile)["75%",]
uppers
```

```{r}
# What proportion of the upper quartiles from the random data are less than the distance upper quartile ?
pval = sum(uppers < distQuartiles["75%"])
pval
```

Q4: Write a function called segments that takes three arguments: from, to, and rep. The function should generate a matrix with two columns, the first column containing from values and the second column containing to values. The argument rep should default to FALSE.

```{r}
# -----  simplest form ----- 
segments = function(from, to, rep = FALSE){
  mat = cbind(from,to)
  return(mat)
}

segments(.1, .2)
```
```{r}
# ----- Add feature: The values in from and to should all be between -1 and 1.           -----
# -----              Values outside that range should be "clamped" to 1 with a warning.  -----

segments = function(from, to, rep = FALSE){
  if(from < -1){
    warning("'from' value(s) less than -1 raised to -1 \n")
    from = -1
  }
  if(from > 1){
    warning("'from' value(s) larger than 1 reduced to 1 \n")
    from = 1
  }
  if(to < -1){
    warning("'to' value(s) less than -1 raised to -1 \n")
    to = -1
  }
  if(to > 1){
    warning("'to' value(s) larger than 1 reduced to 1 \n")
    to = 1
  }
  mat = cbind(from,to)
  return(mat)
}

segments(.1, 1.2)
segments(-1.2, .1)
```
```{r}
# ----- Add feature: The values in from and to can be vectors and they should both "recycle" to the longest length. -----

segments = function(from, to, rep = FALSE){
  # check the validity of from/to value, 
  # convert all invalid values less than -1 to -1
  # convert all invalid values larger than 1 to 1
  if(any(from < -1)){
    warning("'from' value(s) less than -1 raised to -1 \n")
    from = -1
  }
  if(any(from > 1)){
    warning("'from' value(s) larger than 1 reduced to 1 \n")
    from = 1
  }
  if(any(to < -1)){
    warning("'to' value(s) less than -1 raised to -1 \n")
    to = -1
  }
  if(any(to > 1)){
    warning("'to' value(s) larger than 1 reduced to 1 \n")
    to = 1
  }
  
  mat = cbind(from,to)
  return(mat)
}

segments(c(.1, .2), c(.3, .4))
segments(1:4/10, .9)
```
```{r}
# ----- Add feature: Sort rows in order from smallest from value to largest from value.

segments = function(from, to, rep = FALSE){
  # check the validity of from/to value, 
  # convert all invalid values less than -1 to -1
  # convert all invalid values larger than 1 to 1
  if(any(from < -1)){
    warning("'from' value(s) less than -1 raised to -1 \n")
    from = -1
  }
  if(any(from > 1)){
    warning("'from' value(s) larger than 1 reduced to 1 \n")
    from = 1
  }
  if(any(to < -1)){
    warning("'to' value(s) less than -1 raised to -1 \n")
    to = -1
  }
  if(any(to > 1)){
    warning("'to' value(s) larger than 1 reduced to 1 \n")
    to = 1
  }
  
  mat = cbind(from,to)
  
  # sort rows in order from smallest from value to largest from value.
  mat = mat[order(mat[,1]), ]

  return(mat)
}

segments(4:1/10, .9)

```

```{r}
# ----- Add feature: Treat any negative values in from/to as values measuring backward from 1.

segments = function(from, to, rep = FALSE){
  # check the validity of from/to value, 
  # convert all invalid values less than -1 to -1
  # convert all invalid values larger than 1 to 1
  if(any(from < -1)){
    warning("'from' value(s) less than -1 raised to -1 \n")
    from = -1
  }
  if(any(from > 1)){
    warning("'from' value(s) larger than 1 reduced to 1 \n")
    from = 1
  }
  if(any(to < -1)){
    warning("'to' value(s) less than -1 raised to -1 \n")
    to = -1
  }
  if(any(to > 1)){
    warning("'to' value(s) larger than 1 reduced to 1 \n")
    to = 1
  }
  
  # treat any negative values in from/to as values measuring backward from 1.
  from[from < 0] = from[from < 0] + 1.0
  to[to < 0] = to[to < 0] + 1.0
  
  mat = cbind(from,to)
  
  # sort rows in order from smallest from value to largest from value.
  mat = mat[order(mat[,1]), ]
  
  return(mat)
}

segments(.1, -.1)
```
```{r}
segments(c(.1, .2), c(.7, -.2))
```

```{r}
# ----- Add feature: If any from values are larger than the corresponding to values, they should be swapped. -----

segments = function(from, to, rep = FALSE){
  # check the validity of from/to value, 
  # convert all invalid values less than -1 to -1
  # convert all invalid values larger than 1 to 1
  if(any(from < -1)){
    warning("'from' value(s) less than -1 raised to -1 \n")
    from = -1
  }
  if(any(from > 1)){
    warning("'from' value(s) larger than 1 reduced to 1 \n")
    from = 1
  }
  if(any(to < -1)){
    warning("'to' value(s) less than -1 raised to -1 \n")
    to = -1
  }
  if(any(to > 1)){
    warning("'to' value(s) larger than 1 reduced to 1 \n")
    to = 1
  }
  
  # treat any negative values in from/to as values measuring backward from 1.
  from[from < 0] = from[from < 0] + 1.0
  to[to < 0] = to[to < 0] + 1.0
  
  mat = cbind(from,to)
  
  # boolSwap:rows whose from value > to value
  boolSwap = mat[, "from"] > mat[, "to"]
  
  # swap the values of from and to for these rows
  tmptSwap = mat[, "from"][boolSwap] 
  mat[, "from"][boolSwap] = mat[, "to"][boolSwap]
  mat[, "to"][boolSwap] = tmptSwap
  
  # sort rows in order from smallest from value to largest from value.
  mat = mat[order(mat[,1]), ]
  
  return(mat)
}

segments(.9, .1)
```

```{r}

# ----- Add feature: If rep is TRUE, "repeat" the values in the argument until they exceed 1. -----

segments = function(from, to, rep = FALSE){
  # check the validity of from/to value, 
  # convert all invalid values less than -1 to -1
  # convert all invalid values larger than 1 to 1
  if(any(from < -1)){
    warning("'from' value(s) less than -1 raised to -1 \n")
    from = -1
  }
  if(any(from > 1)){
    warning("'from' value(s) larger than 1 reduced to 1 \n")
    from = 1
  }
  if(any(to < -1)){
    warning("'to' value(s) less than -1 raised to -1 \n")
    to = -1
  }
  if(any(to > 1)){
    warning("'to' value(s) larger than 1 reduced to 1 \n")
    to = 1
  }
  
  # treat any negative values in from/to as values measuring backward from 1.
  from[from < 0] = from[from < 0] + 1.0
  to[to < 0] = to[to < 0] + 1.0
  
  mat = cbind(from,to)
  
  # boolSwap:rows whose from value > to value
  boolSwap = mat[, "from"] > mat[, "to"]
  
  # swap the values of from and to for these rows
  tmptSwap = mat[, "from"][boolSwap] 
  mat[, "from"][boolSwap] = mat[, "to"][boolSwap]
  mat[, "to"][boolSwap] = tmptSwap
  
  # sort rows in order from smallest from value to largest from value.
  mat = mat[order(mat[,1]), ]
  
  # If rep is TRUE, "repeat" the values in the argument until they exceed 1.
  if(rep & mat[length(mat)] < 1.0){
    # if only one row, pattern is determined by the diff between from and to
    if(length(mat) == 2)
    {
      step = 2 * (mat[2] - mat[1])
      to = seq(mat[2], to = 1.0, by = step )
      from = seq(mat[1], by = step, length = length(to) )
    }
    # if more than one row, pattern is determined by the last two value in each column
    else{
      # extend column to (expanded part)
      stepTo = mat[nrow(mat), "to"] - mat[nrow(mat)-1, "to"] 
      extendTo = seq(mat[nrow(mat), "to"] + stepTo , to = 1.0, by = stepTo)
      # extend column from (expanded part)
      stepFrom = mat[nrow(mat), "from"] - mat[nrow(mat)-1, "from"]
      extendFrom = seq(mat[nrow(mat), "from"] + stepFrom , by = stepFrom, length = length(extendTo))
      # new  = origin + expanded
      from = c(mat[,"from"], extendFrom)    
      to = c(mat[,"to"], extendTo)
    }
    
    # generate the resulted matrix
    mat = cbind(from,to)
    dimnames(mat) = list(NULL, c("from", "to"))
  }
  
  return(mat)
}

segments(.1, .2, rep = TRUE)
segments(c(.3, .1), .2, rep = TRUE)

```

