---
title: "STATS 782 Assignment 3"
author: "Francis Tang UPI: ftan638"
date: "Due: 22 May 2019"
output: pdf_document
---

1.
An solution would be using arrows() to plot this figure.

```{r}
pdf("Q1.pdf", onefile=TRUE, width=7.2, height=5.5)
plot.new()
plot.window(xlim = c(-100,100), ylim = c(-100, 100))
rect(-50, -70, 50, 70, 
     col = rgb(227, 100, 65, maxColorValue = 255), 
     border = "white")
xl <- -30
xr <- 30 
yb <- -25 
yt <- 25
arrows(xl, yt, xr, yt, code = 3, angle = 125, lwd = 5, 
       length = 0.4, lend = 3, col = "white")
arrows(xl, yb, xr, yb, code = 3, angle = 55,  lwd = 5, 
       length = 0.4, lend = 3, col = "white")
dev.off()
```

\begin{center}
\includegraphics[width=6in]{Q1.pdf}
\end{center}


2.

(a) Mimic the figure.

```{r}
pdf("Q2a.pdf", width = 7.2, height = 5)
plot.new()
plot.window(xlim = c(-3.5,3.5), ylim = c(-1,1))
x1<- seq(-5, -0.862, length=1000)
y1 <- dnorm(x1, mean=0, sd=1)
x2<- seq(-0.862, 5, length=1000)
y2 <- dnorm(x2, mean=0, sd=1)
placex <- c(-1.415,-0.862,0.342)
placey <- c(-0.013,-0.025,-0.03)
par(new = TRUE)
plot(c(x1,x2,placex,0.5), c(y1,y2, placey,-0.06), 
     type = "n", xaxt='n',yaxt='n', ann=FALSE, bty ="n")
clip(-3.7, 3.7, -1, 1)
lines(x1, y1, lwd=2, col = "darkgreen")
lines(x2, y2, lwd=2, col = "blue")
abline(h = 0, lwd = 1.5, col = "blue", lend = 1)
p = dnorm(-0.862, mean = 0, sd = 1.1)
lines(rep(-0.862, 1000), seq(0, p, length = 1000),
      lty = "dashed", lwd = 2, col = "purple")
tricol <- c("darkgreen","purple","blue")
label <- c(expression("c"[1]), 
           expression(paste(mu, "(", omega, " = 0.1)")), 
           expression("c"[2]))
points(placex, placey, pch = c(2, 17, 2), 
       xaxt = 'n', yaxt = 'n', ann = FALSE, 
       col = tricol, cex = c(1, 3, 4), lwd = c(1, 1.5, 1))
mtext(label, side = 1, line = c(-3.4, -0.3, -3.4), 
      at = placex, col = tricol,  cex = 1.5)
dev.off()
```

\begin{center}
\includegraphics[width=4.5in]{Q2a.pdf}
\end{center}

(b) Improve the figure: the original figure above shows a high contrast in colours, for example, the blue and purple are too bright. To solve this, I chose to use some more mild colours.

```{r}
pdf("Q2b.pdf", onefile=TRUE, width=7, height=5)
plot.new()
plot.window(xlim = c(-4, 4), ylim = c(-1,1))
x1<- seq(-5, -0.862, length=1000)
y1 <- dnorm(x1, mean=0, sd=1)
x2<- seq(-0.862, 5, length=1000)
y2 <- dnorm(x2, mean=0, sd=1)
placex <- c(-1.415,-0.862,0.342)
placey <- c(-0.013,-0.025,-0.03)
par(new = TRUE)
plot(c(x1,x2,placex,0.5), c(y1,y2, placey,-0.06), 
     type = "n", xaxt='n', yaxt='n', ann=FALSE, bty ="n")
clip(-3.7, 3.7, -1, 1)
lines(x1, y1, lwd=2, col = "firebrick3")
lines(x2, y2, lwd=2, col = "deepskyblue3")
lines(c(-3.5,3.5), c(0, 0), lwd = 1.5, 
      col = "black", lend = 1)
#abline(h = 0, lwd = 2, col = "dodgerblue3", lend = 1)
p = dnorm(-0.862, mean = 0, sd = 1.1)
lines(rep(-0.862, 1000), seq(0, p, length = 1000),
      lty = "dashed", lwd = 2, col = "forestgreen")
tricol <- c("firebrick3","forestgreen","deepskyblue3")
label <- c(expression("c"[1]), expression(paste(mu, "(", omega, " = 0.1)")), 
           expression("c"[2]))
points(placex, placey, pch = c(2, 17, 2), xaxt = 'n', yaxt = 'n', ann = FALSE, 
       col = tricol, cex = c(1, 3, 4), lwd = c(1, 1.5, 1.5))
mtext(label, side = 1, line = c(-3.4, -0.3, -3.4), 
      at = placex, col = tricol, cex = 1.5)
dev.off()
```

\begin{center}
\includegraphics[width=4.5in]{Q2b.pdf}
\end{center}

3. Mimic the Archimedean Spiral.

```{r}
pdf("Q3.pdf", onefile=TRUE, width=7, height=5)
par(pty="s", bty="n", mar = c(5,5,5,5))
plot.new()
plot.window(xlim = c(-6, 6), ylim = c(-6, 6))

# draw the first Archimedean spiral
a1 <-  0.3
theta1 <- seq(0, 6 * pi, len = 400)

r1 <- a1 * theta1
x1 <- r1 * cos(theta1)
y1 <- r1 * sin(theta1)

# draw the second Archimedean spiral
a2 <- 0.36
theta2 <- seq(0, 5 * pi , len = 400)

r2 <- a2 * theta2
x2 <- r2 * cos(theta2)
y2 <- r2 * sin(theta2)

# draw the red part of the circle
theta3 <- seq(0, pi , 0.01)
r3 <- (1.8 * pi)
x3 <- r3 * cos(theta3)
y3 <- r3 * sin(theta3)

# draw the black part of the circle
theta4 <- seq(pi, 2*pi , 0.01)
r4 <- (1.8 * pi)
x4 <- r4 * cos(theta4)
y4 <- r4 * sin(theta4)

# start to fill the colour in
tx2 <- x2[length(x2) : 1]
ty2 <- y2[length(y2) : 1]
polygon(c(x1, x3, tx2), c(y1, y3, ty2), 
        col = rgb(143, 188, 143, maxColorValue = 255),
        border = NA)

# finish the lines
lines(x1, y1, lwd = 1.3, col = "red")
lines(x3, y3, lwd = 1, col = "black")
lines(x2, y2, lwd = 1.3, col = "red")
lines(x4, y4, lwd = 1.3, col = "red")

dev.off()
```

\begin{center}
\includegraphics[width=5in]{Q3.pdf}
\end{center}

4.

(a) Check integrity.

```{r}
# inport dataset
datasyd <- read.csv("~/Desktop/STATS 782/datasyd.csv")
```

Now we check the annual means column by computing the means over all months.

```{r}
# calculate means for Jan to Dec for each year
rowMeans(datasyd[ ,4:16], na.rm = TRUE)
MeanAnn <- mean(datasyd$Annual, na.rm = TRUE)
```

```{r}
# display the annual column
datasyd$Annual
```

From the result we got above, the annual means over 161 years match the annual means column, which proves the integrity of the data.

(b) Mimic the figure from 1910 to 2018.

```{r}
# pick data from 1910 first
datasyd1910 <- datasyd[-c(1:51, 161), ]

# calculate annual mean from 1061-1990 data
datasyd1961 <- datasyd[c(103:132), ]
MeanAnn1961 <- mean(datasyd1961$Annual, na.rm = TRUE)
datasyd1910$Diff <- datasyd1910$Annual - MeanAnn1961
```

```{r}
# prepare the whole all year's data from question (c)
datasyd1859 <- datasyd[-c(161), ]
MeanAnn1859 <- mean(datasyd1859$Annual, na.rm = TRUE)
datasyd1859$Diff <- datasyd1859$Annual - MeanAnn1961
```

The range of the temperature change is difference from the BBC figure, so I added more ticks for the y-axis.

```{r}
#leave margin for the plot
pdf("Q4a.pdf", onefile=TRUE, width=7, height=5)
par(bty = "n", mar = c(4,5,5,2))
plot.new()
plot.window(xlim = range(datasyd1910$Year), 
            ylim = range(datasyd1910$Diff))
par(new = TRUE)

# x and y axis ticks
xticks1 <- c(datasyd1910$Year[seq(1, 109, by = 18)])
yticks1 <- c(-1.5, -1, -0.5, 0, 0.5, 1, 1.5)

# draw grid in the behind
drawGrid1 <- function(){
  abline(v = xticks1, col = "gray94")
  abline(h = yticks1, col = "gray94")
}

# plot the bars
plot(datasyd1910$Year, datasyd1910$Diff, type = "h", 
     col = ifelse(datasyd1910$Diff < 0, 
                  "deepskyblue4", "firebrick3"), 
     lwd = 3.5, lend = 3, xaxt = 'n', yaxt = 'n', 
     ylab ='', xlab ='', 
     panel.first = drawGrid1())

# draw the axis and lines
axis(1, pos = -1.4, xticks1, cex.axis = 0.9, 
     las = 0, font = 1, tick = FALSE)
abline(h = 0, lwd = 1, col = "black", lend = 1)
lines(c(1897, 2022), c(-2.25, -2.25), lwd = 0.8, 
      col = "black", lend = 1, xpd = TRUE)

# display the text
mtext(yticks1, side = 2, at = yticks1, line = 2, 
      col = 'black', cex = 0.9, adj = 0, las = 2)
mtext(c("Note: Average is calculated from 1961-1990 data"), 
      side = 1,at = 1897, line = 1.5, 
      col = 'black', cex = 0.7, adj = 0)
mtext(c("Source: Australian Government Bureau of Meteorology"), 
      side = 1,at = 1897, line = 2.5, 
      col = 'black', cex = 0.8,adj = 0)
mtext(c("Annual mean temperature above or below average (\u00B0C)"), 
      side = 3,at = 1897, line = 1.5, 
      col = 'black', cex = 1.2,adj = 0)
mtext(c(expression(paste(bold("Australia has been getting warmer")))), 
      side = 3,at = 1897, line = 2.5, 
      col = 'black', cex = 1.5,adj = 0)
dev.off()
```

\begin{center}
\includegraphics[width=5in]{Q4a.pdf}
\end{center}

(c) Add all years data and interpret the figure.

Now we add all years data since 1859, the plot gets wider so I decreased the bar width.

```{r}
pdf("Q4b.pdf", onefile=TRUE, width=7, height=5)
par(bty = "n", mar = c(4,5,5,2))
plot.new()
plot.window(xlim = range(datasyd1859$Year), 
            ylim = range(datasyd1859$Diff))
par(new = TRUE)
xticks <- c(datasyd1859$Year[seq(1, 160, by = 20)], 2018)
yticks <- c(-2, -1.5, -1, -0.5, 0, 0.5, 1, 1.5)
drawGrid <- function(){
  abline(v = xticks, col = "gray94")
  abline(h = yticks, col = "gray94")
}
plot(datasyd1859$Year, datasyd1859$Diff, type = "h", 
     col = ifelse(datasyd1859$Diff < 0, 
                  "deepskyblue4", "firebrick3"), 
     lwd = 2.3, lend = 3, xaxt = 'n', yaxt = 'n', 
     ylab ='', xlab ='', 
     panel.first = drawGrid())
axis(1, pos = -1.8, xticks,
     cex.axis = 0.9, las = 0, font = 1, tick = FALSE)
abline(h = 0, lwd = 1, col = "black", lend = 1)
lines(c(1840, 2022), c(-2.96, -2.96), lwd = 0.8, 
      col = "black", lend = 1, xpd = TRUE)
mtext(yticks, side = 2, at = yticks, line = 2, 
      col = 'black', cex = 0.9, adj = 0, las = 2)
mtext(c("Note: Average is calculated from 1961-1990 data"), 
      side = 1,at = 1840, line = 2, 
      col = 'black', cex = 0.7, adj = 0)
mtext(c("Source: Australian Government Bureau of Meteorology"), 
      side = 1,at = 1840, line = 3, 
      col = 'black', cex = 0.8,adj = 0)
mtext(c("Annual mean temperature above or below average (\u00B0C)"), 
      side = 3,at = 1840, line = 1.5, 
      col = 'black', cex = 1.2,adj = 0)
mtext(c(expression(paste(bold("Australia has been getting warmer")))), 
      side = 3,at = 1840, line = 2.5, 
      col = 'black', cex = 1.5,adj = 0)
dev.off()
```

\begin{center}
\includegraphics[width=5in]{Q4b.pdf}
\end{center}

It is not sufficient enough to see the trend and information behind the figure, so I fit a simple linear regression based on the data and the scatter plot is displayed below with the regression line.

```{r}
pdf("Q4c.pdf", onefile=TRUE, width=7, height=5)
par(mar = c(4,5,4,2))
plot(datasyd1859$Year, datasyd1859$Diff, pch = 21, 
     col = ifelse(datasyd1859$Year < 2000, "deepskyblue4", "firebrick3"), 
     xlab = "Year", ylab = "Temperature Difference")
title("Annual Mean Temperature Above or Below Average (\u00B0C)", line = 2)
abline(lm(formula = Diff ~ Year, data = datasyd1859), lwd = 2, col = "gray47")
dev.off()
```

\begin{center}
\includegraphics[width=5in]{Q4c.pdf}
\end{center}

From the two plots we generated above, we can conclude the following statements:

1. The temperature of Australia keeps arising during the last 150 years.
2. Australia gets warm more rapidly since 2000, which has been shown on the scatter plot, the slope of temperature increase gets larger and the scatters are even above the regression line.
3. Generally speaking, the data from 1859 til now can roughly fit a simple linear regression.

(d) Predict temperature in 2030 and give 95% prediction interval.

```{r}
# first we fit a simple linear regression model
TempDiff.lm = lm(formula = Diff ~ Year, data = datasyd1859)
# check the summary data
summary(TempDiff.lm)
```

```{r}
# use ANOVA
anova(TempDiff.lm)
```

```{r}
# use diagnostic plots
par(mfrow=c(2,2))
plot(TempDiff.lm)
```


After fit a linear regression model on the data, we can see that "Year" regressor shows as a very significant regressor to the model and the diagnostic plots prove it as well. So there exist a significant linear regression between year and temperature change.

The next step is to use the model to predict temperature in 2030. First we are going to predict the time difference then add it into the mean of 1961-1990 data.

```{r}
# use the linear model to make prediction for 2030
pred2030 <- predict(TempDiff.lm, newdata = data.frame(Year = 2030), 
        interval="prediction", level = 0.95) 
temp2030 = MeanAnn1961 + pred2030
temp2030
```

The data above shows us the prediction of 2030 temperature is 22.98 degrees in Celcius. And the 95% prediction interval is between (22.075, 23.883).

(e) Bonus

Here we calculate the probablity of greater than 6.5 degrees in the future:

```{r}
1 - pexp(6.5, rate = (1/2.17), lower.tail = TRUE, log.p = FALSE)
```

So we can see that the temperature is 2.17 at the year, substitute 2.17 into the linear model we fit before:

```{r}
yearPredict <- (2.17 + 2.638e+01) / 1.340e-02
yearPredict - 2100
```

The answer is 30.6 years.




